---
title: "Country portfolio"
author: "Juan Camilo Castillo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Country Portfolio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Finding the optimal portfolio for a country

```{r setup}
library(vaccineEarlyInvest)
```

## Simple case: Peru

The main function that finds the optimal portfolio for a price taking country is `portfolioPriceTaker()`. In order to illustrate its basic usage, let's find the optimal portfolio for Peru. It has 31.99 million people, its GDP per capita is $6,710, 13.1% of its population is high risk, and the World Bank projects the Covid-19 pandemic will cause a 26.9% cumulative GDP loss over 2020 and 2021. The main arguments to `portfolioPriceTaker()` are these five numbers, the price at which manufacturing capacity is sold in the market (let's assume it's $4 per course / year), as well as the path to a .csv file with data on vaccine candidates.

```{r, fig.width=5, fig.height=3.5}
population <- 31.99
gdp_pc <- 6.71
fracHighRisk <- 0.131
loss2yr <- 0.269
price <- 4

portfolio <- portfolioPriceTaker(population, gdp_pc, fracHighRisk, loss2yr, price,
                                 candidateFile="../inst/extdata/vaccinesSummary.csv")
```

The output from this function is a list that includes the main information about the optimal portfolio:

```{r, fig.width=5, fig.height=3.5}
portfolio$expBenefits # In billion dollars
portfolio$cost # In billion dollars
portfolio$totCapacity # In million courses / month

# Sorting capacities in descending order
capacities <- sort(portfolio$capacities, decreasing=T)

library(ggplot2)
theme_set(theme_bw(base_size=12))
ggplot() + geom_bar(aes(1:length(capacities), capacities), stat="identity",
                    fill="black", color="black", alpha=0.5) +
  labs(x="Candidate", y="Capacity (mil. courses / month)", 
       title="Capacity per cand. at $4 per course / month, Peru")
```
