---
title: "Country portfolio"
author: "Juan Camilo Castillo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Country Portfolio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Finding the optimal portfolio for a country

```{r setup}
# Loading packages and setting up options
library(vaccineEarlyInvest)
library(ggplot2)
theme_set(theme_bw(base_size=12))
```

## Simple case: Peru

The main function that finds the optimal portfolio for a price taking country is `portfolioPriceTaker()`. In order to illustrate its basic usage, let's find the optimal portfolio for Peru. It has 31.99 million people, its GDP per capita is $6,710, 13.1% of its population is high risk, and the World Bank projects the Covid-19 pandemic will cause a 26.9% cumulative GDP loss over 2020 and 2021. The main arguments to `portfolioPriceTaker()` are these five numbers, the price at which manufacturing capacity is sold in the market (let's assume it's $4 per course / year), as well as the path to a .csv file with data on vaccine candidates.

```{r, fig.width=5, fig.height=3.5}
population <- 31.99
gdp_pc <- 6.71
frac_high_risk <- 0.131
loss2yr <- 0.269
price <- 4

portfolioPeru <- portfolioPriceTaker(population=population, gdp_pc=gdp_pc, frac_high_risk=frac_high_risk, 
                                     loss2yr=loss2yr, price=price,
                                 candidateFile="../inst/extdata/vaccinesSummary.csv")
```

Note that the names of all six arguments must be entered explicitly when calling `portfolioPriceTaker()`. The output from this function is a list that includes the main information about the optimal portfolio:

```{r, fig.width=5, fig.height=3.5}
portfolioPeru$expBenefits # In billion dollars
portfolioPeru$cost # In billion dollars
portfolioPeru$totCapacity # In million courses / month

# Sorting capacities in descending order
capacities <- sort(portfolioPeru$capacities, decreasing=T)

ggplot() + geom_bar(aes(1:length(capacities), capacities), stat="identity",
                    fill="black", color="black", alpha=0.5) +
  labs(x="Candidate", y="Capacity (mil. courses / month)", 
       title="Capacity per cand. at $4 per course / month, Peru")
```

The function `demandPriceTaker()` finds the whole demand curve for a country for different prices. It is much more efficient than calling `portfolioPriceTaker()` multiple times at different prices. Here is an example of how to call that function and plot total demand for manufacturing capacity for Peru:

```{r, fig.width=5, fig.height=3.5}
prices <- 1:30 # Vector of prices at which demand will be evaluated

demandPeru <- demandPriceTaker(population=population, gdp_pc=gdp_pc, frac_high_risk=frac_high_risk, 
                                     loss2yr=loss2yr, prices=prices,
                                 candidateFile="../inst/extdata/vaccinesSummary.csv")

ggplot() + geom_line(aes(totalCapacity, price), demandPeru$optimizations) + 
  labs(x="Capacity (mil. courses / month)", y="Price ($ per course / year)", 
       title="Demand curve, Peru")
```


## Advanced options

The previous simple analysis uses the default parameters of the model. Those default parameters can be modified by passing an object of the `Parameters` class. In that case, the input arguments `population`, `gdp_pc`, `frac_high_risk`, and `loss2yr` passed to `portfolioPriceTaker()` and `demandPriceTaker()` are overriden. Instead, they should be passed as arguments to the constructor of the `Parameters` class.

As an example, consider


